---
title: "Sex composition of infant filicide relative to homicide generally"
author: "ChoJin"
date: "2026-09-06"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = file.path(dirname(inputFile), "output")) })
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
options(stringsAsFactors = FALSE, scipen = 6)
```

## Descriptive question

> How unusual is the sex composition of infant filicide relative to homicide
> generally?

This first analysis calculates three descriptive quantities for each group:

1. **Proportion female** among offender records with known sex:
   $F/(F+M)$.
2. **Female:male odds**: $F/M$.
3. **Odds ratio relative to general homicide**:
   $(F_{subtype}/M_{subtype})/(F_{general}/M_{general})$.

The analysis pools the latest ten complete FBI data years, 2016--2025, before
calculating these quantities. Pooling the counts answers the question about the
sex composition of the observed offender records across the period. It does
not average the ten annual percentages.

## Analytic definitions

- The data are the FBI Supplementary Homicide Report (SHR) annual master files
  stored in `data/raw/fbi_shr/`.
- General homicide includes records coded murder or nonnegligent manslaughter.
  SHR circumstance codes 80 and 81, which identify justifiable homicide, are
  excluded.
- A filicide group requires a first-listed victim from birth through age 17
  and a victim-to-offender relationship of son (`SO`), daughter (`DA`),
  stepson (`SS`), or stepdaughter (`SD`). Thus, parent and stepparent records
  are included. Infant filicide is the birth-through-364-day subset.
- The SHR age code `NB` means birth through 6 days; `BB` means 7 through 364
  days. The combined infant category contains both. Numeric ages 01--17 are
  grouped as 1--4, 5--9, 10--14, and 15--17 years. The table also includes an
  all-filicide row covering birth through age 17.
- The analytic unit is a represented **offender record**, not an incident or a
  person tracked across incidents. An incident with multiple recorded
  offenders can contribute multiple offender records.
- Offender records coded male or female form the denominator for the requested
  sex-composition measures. Unknown and not-specified sex records are retained
  as a separate count but excluded from the proportion and odds calculations.
- In the SHR fixed-width layout, relationships are available for the
  first-listed victim only. Consequently, the filicide groups do not capture a
  child listed only as an additional victim in a multi-victim incident.

## Read and parse the FBI SHR files

The archive parsing code uses base R plus `knitr`, which is already used to
render the document. The inferential section later uses `sandwich` for
cluster-robust covariance estimation. No Python-generated analytic table is
imported.

```{r parameters}
analysis_start_year <- 2016L
analysis_end_year <- 2025L
analysis_years <- analysis_start_year:analysis_end_year

shr_directory <- file.path("data", "raw", "fbi_shr")
output_directory <- "output"
output_csv <- file.path(
  output_directory,
  sprintf(
    "homicide_subtype_sex_comparison_%d_%d.csv",
    analysis_start_year,
    analysis_end_year
  )
)

record_length <- 268L
maximum_additional_offenders_in_layout <- 10L
filicide_relationship_codes <- c("SO", "DA", "SS", "SD")
sex_levels <- c("male", "female", "unknown/not specified")
filicide_detailed_subtype_ids <- c(
  "filicide_birth_through_6_days",
  "filicide_days_7_through_364",
  "filicide_years_1_through_4",
  "filicide_years_5_through_9",
  "filicide_years_10_through_14",
  "filicide_years_15_through_17"
)
model_group_ids <- c("other_homicide", filicide_detailed_subtype_ids)
```

```{r parser-functions}
read_shr_archive <- function(year, directory = shr_directory) {
  archive <- file.path(directory, sprintf("shr-%d.zip", year))
  if (!file.exists(archive)) {
    stop(
      sprintf(
        "Missing %s. The FBI SHR archive must be present before knitting.",
        archive
      )
    )
  }

  members <- utils::unzip(archive, list = TRUE)
  members <- members[!grepl("/$", members$Name), , drop = FALSE]
  if (nrow(members) != 1L) {
    stop(sprintf("Expected one data member in %s; found %d.", archive, nrow(members)))
  }

  connection <- unz(archive, members$Name[[1L]], open = "rt")
  on.exit(close(connection), add = TRUE)
  records <- readLines(connection, warn = FALSE, encoding = "latin1")

  # Legacy DAT files can end with a one-character DOS end-of-file record.
  records <- records[nchar(records, type = "chars") >= 90L]
  lengths <- nchar(records, type = "chars")
  if (any(lengths > record_length)) {
    stop(sprintf("At least one record in %s exceeds %d characters.", archive, record_length))
  }

  # Some official mid-2000s files physically omit trailing blank fields.
  # Right-padding restores the documented logical fixed-width layout.
  paste0(records, strrep(" ", record_length - lengths))
}

normalize_sex <- function(raw_code) {
  ifelse(
    raw_code == "M",
    "male",
    ifelse(raw_code == "F", "female", "unknown/not specified")
  )
}

classify_filicide_age <- function(raw_code) {
  if (raw_code == "NB") {
    return("filicide_birth_through_6_days")
  }
  if (raw_code == "BB") {
    return("filicide_days_7_through_364")
  }

  age <- suppressWarnings(as.integer(raw_code))
  if (is.na(age) || age < 1L || age > 17L) {
    return(NA_character_)
  }
  if (age <= 4L) {
    return("filicide_years_1_through_4")
  }
  if (age <= 9L) {
    return("filicide_years_5_through_9")
  }
  if (age <= 14L) {
    return("filicide_years_10_through_14")
  }
  "filicide_years_15_through_17"
}

extract_offender_segments <- function(record) {
  reported_additional <- suppressWarnings(as.integer(substr(record, 96L, 98L)))
  if (is.na(reported_additional)) {
    stop("Encountered an invalid additional-offender count.")
  }
  represented_additional <- min(
    reported_additional,
    maximum_additional_offenders_in_layout
  )

  segments <- substr(record, 81L, 92L)
  if (represented_additional > 0L) {
    starts <- 149L + (seq_len(represented_additional) - 1L) * 12L
    segments <- c(
      segments,
      vapply(starts, function(start) substr(record, start, start + 11L), character(1L))
    )
  }
  segments
}

empty_sex_counts <- function() {
  stats::setNames(integer(length(sex_levels)), sex_levels)
}

parse_shr_year <- function(year) {
  records <- read_shr_archive(year)
  # Position 71: A = murder and nonnegligent manslaughter.
  records <- records[substr(records, 71L, 71L) == "A"]

  general_counts <- empty_sex_counts()
  filicide_subtype_ids <- c(
    "filicide_under_18",
    "filicide_under_1_year",
    filicide_detailed_subtype_ids
  )
  filicide_counts <- stats::setNames(
    lapply(filicide_subtype_ids, function(subtype_id) empty_sex_counts()),
    filicide_subtype_ids
  )
  offender_record_chunks <- vector("list", length(records))

  for (record_index in seq_along(records)) {
    record <- records[[record_index]]
    segments <- extract_offender_segments(record)

    # Within a 12-character offender segment, positions 10-11 are the
    # circumstance. Codes 80/81 are justifiable homicide.
    circumstances <- substr(segments, 10L, 11L)
    segments <- segments[!circumstances %in% c("80", "81")]
    if (length(segments) == 0L) {
      next
    }

    offender_sex <- normalize_sex(substr(segments, 3L, 3L))
    general_counts <- general_counts + table(
      factor(offender_sex, levels = sex_levels)
    )

    first_victim_age <- substr(record, 76L, 77L)
    detailed_subtype_id <- classify_filicide_age(first_victim_age)
    # Within a 12-character offender segment, positions 8-9 are the
    # relationship of the first-listed victim to that offender.
    relationships <- substr(segments, 8L, 9L)
    qualifying <-
      !is.na(detailed_subtype_id) &
      relationships %in% filicide_relationship_codes

    homicide_group <- rep("other_homicide", length(segments))
    homicide_group[qualifying] <- detailed_subtype_id
    offender_record_chunks[[record_index]] <- data.frame(
      year = rep.int(year, length(segments)),
      source_record_id = sprintf("%d-%07d", year, record_index),
      homicide_group = homicide_group,
      offender_sex = offender_sex,
      stringsAsFactors = FALSE
    )

    if (any(qualifying)) {
      qualifying_sex <- offender_sex[qualifying]
      qualifying_counts <- table(factor(qualifying_sex, levels = sex_levels))
      filicide_counts[["filicide_under_18"]] <-
        filicide_counts[["filicide_under_18"]] + qualifying_counts
      filicide_counts[[detailed_subtype_id]] <-
        filicide_counts[[detailed_subtype_id]] + qualifying_counts

      if (detailed_subtype_id %in% filicide_detailed_subtype_ids[1:2]) {
        filicide_counts[["filicide_under_1_year"]] <-
          filicide_counts[["filicide_under_1_year"]] + qualifying_counts
      }
    }
  }

  list(
    counts = c(list(general_homicide = general_counts), filicide_counts),
    offender_records = do.call(rbind, offender_record_chunks)
  )
}
```

```{r parse-years}
annual_parsed_data <- lapply(analysis_years, parse_shr_year)
names(annual_parsed_data) <- analysis_years

annual_counts <- lapply(annual_parsed_data, function(year_data) year_data$counts)
offender_records <- do.call(
  rbind,
  lapply(annual_parsed_data, function(year_data) year_data$offender_records)
)
rownames(offender_records) <- NULL

subtype_ids <- names(annual_counts[[1L]])
pooled_counts <- lapply(
  subtype_ids,
  function(subtype_id) {
    Reduce(
      `+`,
      lapply(annual_counts, function(year_data) year_data[[subtype_id]])
    )
  }
)
names(pooled_counts) <- subtype_ids

# Internal consistency checks: the detailed groups must sum exactly to their
# two combined categories.
stopifnot(
  nrow(offender_records) == sum(pooled_counts$general_homicide),
  all(offender_records$homicide_group %in% model_group_ids),
  all(
    pooled_counts$filicide_under_1_year ==
      pooled_counts$filicide_birth_through_6_days +
        pooled_counts$filicide_days_7_through_364
  ),
  all(
    pooled_counts$filicide_under_18 ==
      Reduce(`+`, pooled_counts[filicide_detailed_subtype_ids])
  )
)
```

## Compute the descriptive comparison

```{r compute-comparison}
subtype_labels <- c(
  general_homicide = "General homicide (reference)",
  filicide_under_18 = "Filicide, birth-17 years",
  filicide_under_1_year = "Infant filicide, birth-364 days",
  filicide_birth_through_6_days = "Filicide, birth-6 days",
  filicide_days_7_through_364 = "Infant filicide, 7-364 days",
  filicide_years_1_through_4 = "Filicide, 1-4 years",
  filicide_years_5_through_9 = "Filicide, 5-9 years",
  filicide_years_10_through_14 = "Filicide, 10-14 years",
  filicide_years_15_through_17 = "Filicide, 15-17 years"
)

comparison <- data.frame(
  analysis_start_year = analysis_start_year,
  analysis_end_year = analysis_end_year,
  homicide_subtype = unname(subtype_labels[subtype_ids]),
  male_offender_records = vapply(
    pooled_counts,
    function(counts) unname(counts[["male"]]),
    integer(1L)
  ),
  female_offender_records = vapply(
    pooled_counts,
    function(counts) unname(counts[["female"]]),
    integer(1L)
  ),
  unknown_or_not_specified_offender_records = vapply(
    pooled_counts,
    function(counts) unname(counts[["unknown/not specified"]]),
    integer(1L)
  )
)

comparison$known_sex_offender_records <-
  comparison$male_offender_records + comparison$female_offender_records
comparison$total_offender_records <-
  comparison$known_sex_offender_records +
  comparison$unknown_or_not_specified_offender_records
comparison$female_proportion_known_sex <-
  comparison$female_offender_records / comparison$known_sex_offender_records
comparison$female_to_male_odds <-
  comparison$female_offender_records / comparison$male_offender_records

general_homicide_odds <- comparison$female_to_male_odds[
  subtype_ids == "general_homicide"
]
comparison$odds_ratio_vs_general_homicide <-
  comparison$female_to_male_odds / general_homicide_odds

stopifnot(
  nrow(comparison) == length(subtype_labels),
  length(general_homicide_odds) == 1L,
  comparison$odds_ratio_vs_general_homicide[[1L]] == 1
)

female_proportion_by_subtype <- stats::setNames(
  comparison$female_proportion_known_sex,
  subtype_ids
)
female_to_male_odds_by_subtype <- stats::setNames(
  comparison$female_to_male_odds,
  subtype_ids
)
odds_ratio_by_subtype <- stats::setNames(
  comparison$odds_ratio_vs_general_homicide,
  subtype_ids
)

dir.create(output_directory, recursive = TRUE, showWarnings = FALSE)
utils::write.csv(comparison, output_csv, row.names = FALSE, na = "")
```

```{r comparison-table, echo=FALSE}
display_table <- data.frame(
  `Homicide subtype` = comparison$homicide_subtype,
  Male = format(comparison$male_offender_records, big.mark = ","),
  Female = format(comparison$female_offender_records, big.mark = ","),
  `Unknown / not specified` = format(
    comparison$unknown_or_not_specified_offender_records,
    big.mark = ","
  ),
  `Female proportion` = sprintf(
    "%.1f%%",
    100 * comparison$female_proportion_known_sex
  ),
  `Female:male odds` = sprintf("%.3f", comparison$female_to_male_odds),
  `OR vs. general homicide` = sprintf(
    "%.3f",
    comparison$odds_ratio_vs_general_homicide
  ),
  check.names = FALSE
)

knitr::kable(
  display_table,
  align = c("l", "r", "r", "r", "r", "r", "r"),
  caption = sprintf(
    "Offender-sex composition in FBI SHR records, %d-%d",
    analysis_start_year,
    analysis_end_year
  )
)
```

## Descriptive result

Across 2016--2025, the female proportion among general-homicide offender
records with known sex was
`r sprintf("%.1f%%", 100 * female_proportion_by_subtype[["general_homicide"]])`.
For infant filicide as a whole, it was
`r sprintf("%.1f%%", 100 * female_proportion_by_subtype[["filicide_under_1_year"]])`.
The corresponding female:male odds were
`r sprintf("%.3f", female_to_male_odds_by_subtype[["general_homicide"]])` and
`r sprintf("%.3f", female_to_male_odds_by_subtype[["filicide_under_1_year"]])`, respectively.
Thus, the female:male odds in recorded infant filicide were
`r sprintf("%.2f", odds_ratio_by_subtype[["filicide_under_1_year"]])` times the
female:male odds in homicide generally.

The infant period is not compositionally uniform. The odds ratio was
`r sprintf("%.2f", odds_ratio_by_subtype[["filicide_birth_through_6_days"]])` for
filicide during the first 6 days after birth and
`r sprintf("%.2f", odds_ratio_by_subtype[["filicide_days_7_through_364"]])` for days
7--364. These are descriptive comparisons of reported offender records, not
causal estimates and not statements about any individual case.

The unrounded table is written during knitting to `r output_csv`.

## Inferential model preparation

Formal comparisons require mutually exclusive groups. The inferential data
therefore use the six detailed filicide age groups and a reference group of
all other homicide offender records. The reference excludes every offender
record assigned to one of the six filicide groups. The combined birth--364-day
and birth--17-year rows remain descriptive rollups and are not entered as
additional model observations.

The outcome is female versus male among records with known sex. Unknown or
not-specified sex is analyzed separately as missing outcome information. Each
source-record identifier groups offender segments represented on the same SHR
fixed-width record; cluster-robust inference below allows those segments to be
correlated.

```{r inferential-data}
model_input_csv <- file.path(
  output_directory,
  sprintf(
    "inferential_model_input_%d_%d.csv",
    analysis_start_year,
    analysis_end_year
  )
)
model_results_csv <- file.path(
  output_directory,
  sprintf(
    "inferential_model_results_%d_%d.csv",
    analysis_start_year,
    analysis_end_year
  )
)
model_diagnostics_csv <- file.path(
  output_directory,
  sprintf(
    "inferential_model_diagnostics_%d_%d.csv",
    analysis_start_year,
    analysis_end_year
  )
)
cell_diagnostics_csv <- file.path(
  output_directory,
  sprintf(
    "inferential_model_cell_diagnostics_%d_%d.csv",
    analysis_start_year,
    analysis_end_year
  )
)
missingness_csv <- file.path(
  output_directory,
  sprintf(
    "inferential_model_missingness_%d_%d.csv",
    analysis_start_year,
    analysis_end_year
  )
)
period_sensitivity_csv <- file.path(
  output_directory,
  sprintf(
    "inferential_model_period_sensitivity_%d_%d.csv",
    analysis_start_year,
    analysis_end_year
  )
)
leave_one_year_out_csv <- file.path(
  output_directory,
  sprintf(
    "inferential_model_leave_one_year_out_%d_%d.csv",
    analysis_start_year,
    analysis_end_year
  )
)

model_group_labels <- c(
  other_homicide = "All other homicide (reference)",
  subtype_labels[filicide_detailed_subtype_ids]
)

offender_records$homicide_group <- factor(
  offender_records$homicide_group,
  levels = model_group_ids
)
offender_records$offender_sex <- factor(
  offender_records$offender_sex,
  levels = sex_levels
)

count_array <- table(
  year = factor(offender_records$year, levels = analysis_years),
  homicide_group = offender_records$homicide_group,
  offender_sex = offender_records$offender_sex
)

# expand.grid varies its first factor fastest, matching the vectorized first
# dimension of each year-by-group slice in count_array.
annual_model_counts <- expand.grid(
  year = analysis_years,
  homicide_group = model_group_ids,
  KEEP.OUT.ATTRS = FALSE,
  stringsAsFactors = FALSE
)
annual_model_counts$homicide_group_label <- unname(
  model_group_labels[annual_model_counts$homicide_group]
)
annual_model_counts$male_offender_records <- as.integer(
  count_array[, , "male"]
)
annual_model_counts$female_offender_records <- as.integer(
  count_array[, , "female"]
)
annual_model_counts$unknown_or_not_specified_offender_records <- as.integer(
  count_array[, , "unknown/not specified"]
)
annual_model_counts$known_sex_offender_records <-
  annual_model_counts$male_offender_records +
  annual_model_counts$female_offender_records
annual_model_counts$total_offender_records <-
  annual_model_counts$known_sex_offender_records +
  annual_model_counts$unknown_or_not_specified_offender_records
annual_model_counts$female_proportion_known_sex <- with(
  annual_model_counts,
  female_offender_records / known_sex_offender_records
)
annual_model_counts$unknown_proportion_all_records <- with(
  annual_model_counts,
  unknown_or_not_specified_offender_records / total_offender_records
)

stopifnot(
  sum(annual_model_counts$total_offender_records) == nrow(offender_records),
  all(annual_model_counts$known_sex_offender_records > 0L)
)

utils::write.csv(annual_model_counts, model_input_csv, row.names = FALSE, na = "")
```

## Primary year-adjusted model

The primary model is a binary logistic regression with categorical homicide
group and categorical data year. Age is deliberately categorical because the
age intervals have unequal widths and the two infant codes are qualitatively
distinct. Standard errors are clustered by SHR source record.

```{r primary-inferential-model}
if (!requireNamespace("sandwich", quietly = TRUE)) {
  stop(
    paste(
      "The inferential section requires the sandwich package.",
      "Install it with install.packages('sandwich') and knit again."
    )
  )
}

known_sex_records <- offender_records[
  offender_records$offender_sex %in% c("male", "female"),
  ,
  drop = FALSE
]
known_sex_records$homicide_group <- factor(
  known_sex_records$homicide_group,
  levels = model_group_ids
)
known_sex_records$is_female <- as.integer(
  known_sex_records$offender_sex == "female"
)
known_sex_records$year_factor <- factor(known_sex_records$year)

primary_model <- stats::glm(
  is_female ~ homicide_group + year_factor,
  family = stats::binomial(),
  data = known_sex_records
)
primary_cluster_vcov <- sandwich::vcovCL(
  primary_model,
  cluster = known_sex_records$source_record_id,
  type = "HC1"
)

robust_wald_test <- function(model, covariance, coefficient_names) {
  beta <- stats::coef(model)[coefficient_names]
  covariance_subset <- covariance[
    coefficient_names,
    coefficient_names,
    drop = FALSE
  ]
  statistic <- drop(crossprod(beta, solve(covariance_subset, beta)))
  degrees_of_freedom <- length(beta)
  c(
    statistic = statistic,
    degrees_of_freedom = degrees_of_freedom,
    p_value = stats::pchisq(
      statistic,
      df = degrees_of_freedom,
      lower.tail = FALSE
    )
  )
}

group_coefficient_names <- paste0(
  "homicide_group",
  filicide_detailed_subtype_ids
)
stopifnot(
  primary_model$converged,
  all(group_coefficient_names %in% names(stats::coef(primary_model)))
)

omnibus_group_test <- robust_wald_test(
  primary_model,
  primary_cluster_vcov,
  group_coefficient_names
)

group_beta <- stats::coef(primary_model)[group_coefficient_names]
group_standard_error <- sqrt(diag(primary_cluster_vcov))[group_coefficient_names]
group_z <- group_beta / group_standard_error
group_p <- 2 * stats::pnorm(abs(group_z), lower.tail = FALSE)

pooled_model_counts <- data.frame(
  homicide_group = model_group_ids,
  homicide_group_label = unname(model_group_labels[model_group_ids]),
  male_offender_records = colSums(count_array[, , "male"]),
  female_offender_records = colSums(count_array[, , "female"]),
  unknown_or_not_specified_offender_records = colSums(
    count_array[, , "unknown/not specified"]
  ),
  row.names = NULL,
  check.names = FALSE
)
pooled_model_counts$known_sex_offender_records <- with(
  pooled_model_counts,
  male_offender_records + female_offender_records
)
pooled_model_counts$total_offender_records <- with(
  pooled_model_counts,
  known_sex_offender_records + unknown_or_not_specified_offender_records
)
pooled_model_counts$female_proportion_known_sex <- with(
  pooled_model_counts,
  female_offender_records / known_sex_offender_records
)
pooled_model_counts$female_to_male_odds <- with(
  pooled_model_counts,
  female_offender_records / male_offender_records
)

wilson_interval <- function(successes, trials, confidence_level = 0.95) {
  z <- stats::qnorm(1 - (1 - confidence_level) / 2)
  proportion <- successes / trials
  denominator <- 1 + z^2 / trials
  center <- (proportion + z^2 / (2 * trials)) / denominator
  half_width <- z * sqrt(
    proportion * (1 - proportion) / trials + z^2 / (4 * trials^2)
  ) / denominator
  c(lower = center - half_width, upper = center + half_width)
}

proportion_intervals <- t(vapply(
  seq_len(nrow(pooled_model_counts)),
  function(row_number) {
    wilson_interval(
      pooled_model_counts$female_offender_records[[row_number]],
      pooled_model_counts$known_sex_offender_records[[row_number]]
    )
  },
  numeric(2L)
))

inferential_results <- pooled_model_counts
inferential_results$female_proportion_ci_lower <- proportion_intervals[, "lower"]
inferential_results$female_proportion_ci_upper <- proportion_intervals[, "upper"]
inferential_results$adjusted_odds_ratio_vs_other_homicide <- c(
  1,
  exp(unname(group_beta))
)
inferential_results$cluster_robust_ci_lower <- c(
  1,
  exp(unname(group_beta - stats::qnorm(0.975) * group_standard_error))
)
inferential_results$cluster_robust_ci_upper <- c(
  1,
  exp(unname(group_beta + stats::qnorm(0.975) * group_standard_error))
)
inferential_results$p_value <- c(NA_real_, unname(group_p))
inferential_results$holm_adjusted_p_value <- c(
  NA_real_,
  stats::p.adjust(unname(group_p), method = "holm")
)

utils::write.csv(inferential_results, model_results_csv, row.names = FALSE, na = "")
```

```{r inferential-results-table, echo=FALSE}
inferential_display <- data.frame(
  Group = inferential_results$homicide_group_label,
  Male = format(inferential_results$male_offender_records, big.mark = ","),
  Female = format(inferential_results$female_offender_records, big.mark = ","),
  `Female proportion (95% Wilson CI)` = sprintf(
    "%.1f%% (%.1f%%, %.1f%%)",
    100 * inferential_results$female_proportion_known_sex,
    100 * inferential_results$female_proportion_ci_lower,
    100 * inferential_results$female_proportion_ci_upper
  ),
  `Adjusted OR (95% cluster-robust CI)` = ifelse(
    inferential_results$homicide_group == "other_homicide",
    "1.000 (reference)",
    sprintf(
      "%.3f (%.3f, %.3f)",
      inferential_results$adjusted_odds_ratio_vs_other_homicide,
      inferential_results$cluster_robust_ci_lower,
      inferential_results$cluster_robust_ci_upper
    )
  ),
  `Holm-adjusted p` = ifelse(
    is.na(inferential_results$holm_adjusted_p_value),
    "--",
    format.pval(
      inferential_results$holm_adjusted_p_value,
      digits = 3,
      eps = 0.001
    )
  ),
  check.names = FALSE
)

knitr::kable(
  inferential_display,
  align = c("l", "r", "r", "l", "l", "r"),
  caption = sprintf(
    paste(
      "Year-adjusted female-offender odds by mutually exclusive homicide",
      "group, %d-%d"
    ),
    analysis_start_year,
    analysis_end_year
  )
)
```

The cluster-robust omnibus Wald test of the six homicide-group coefficients is
$\chi^2(`r as.integer(omnibus_group_test[["degrees_of_freedom"]])`) =$
`r sprintf("%.2f", omnibus_group_test[["statistic"]])`,
$p$ `r format.pval(omnibus_group_test[["p_value"]], digits = 3, eps = 0.001)`.
This tests whether the recorded female:male odds in at least one detailed
filicide group differ from the odds among all other homicide records after
adjustment for data year.

## Suitability checks and model diagnostics

```{r model-diagnostics}
# A grouped version of the same main-effects model supports cell-level
# residual, goodness-of-fit, overdispersion, and influence diagnostics.
annual_model_counts$homicide_group_factor <- factor(
  annual_model_counts$homicide_group,
  levels = model_group_ids
)
annual_model_counts$year_factor <- factor(annual_model_counts$year)
grouped_model <- stats::glm(
  cbind(female_offender_records, male_offender_records) ~
    homicide_group_factor + year_factor,
  family = stats::binomial(),
  data = annual_model_counts
)

pearson_chisq <- sum(stats::residuals(grouped_model, type = "pearson")^2)
pearson_degrees_of_freedom <- stats::df.residual(grouped_model)
pearson_dispersion <- pearson_chisq / pearson_degrees_of_freedom
pearson_gof_p <- stats::pchisq(
  pearson_chisq,
  df = pearson_degrees_of_freedom,
  lower.tail = FALSE
)

cell_diagnostics <- annual_model_counts[c(
  "year",
  "homicide_group",
  "homicide_group_label",
  "male_offender_records",
  "female_offender_records",
  "known_sex_offender_records"
)]
cell_diagnostics$fitted_female_proportion <- stats::fitted(grouped_model)
cell_diagnostics$pearson_residual <- stats::residuals(
  grouped_model,
  type = "pearson"
)
cell_diagnostics$standardized_pearson_residual <- stats::rstandard(
  grouped_model,
  type = "pearson"
)
cell_diagnostics$leverage <- stats::hatvalues(grouped_model)
cell_diagnostics$cooks_distance <- stats::cooks.distance(grouped_model)
influence_threshold <- 4 / nrow(cell_diagnostics)
cell_diagnostics$flag_cooks_distance <-
  cell_diagnostics$cooks_distance > influence_threshold

# Check whether the filicide/general contrast differs across the reporting
# transition by testing all group-by-period interactions jointly.
known_sex_records$transition_period <- factor(
  ifelse(known_sex_records$year <= 2020L, "2016-2020", "2021-2025"),
  levels = c("2016-2020", "2021-2025")
)
transition_model <- stats::glm(
  is_female ~ homicide_group * transition_period,
  family = stats::binomial(),
  data = known_sex_records
)
transition_cluster_vcov <- sandwich::vcovCL(
  transition_model,
  cluster = known_sex_records$source_record_id,
  type = "HC1"
)
transition_interaction_names <- grep(
  "^homicide_group.*:transition_period",
  names(stats::coef(transition_model)),
  value = TRUE
)
transition_interaction_test <- robust_wald_test(
  transition_model,
  transition_cluster_vcov,
  transition_interaction_names
)

estimate_period_specific_odds_ratios <- function(period_label, period_years) {
  period_records <- known_sex_records[
    known_sex_records$year %in% period_years,
    ,
    drop = FALSE
  ]
  period_records$homicide_group <- factor(
    period_records$homicide_group,
    levels = model_group_ids
  )
  period_records$year_factor <- factor(period_records$year)

  period_model <- stats::glm(
    is_female ~ homicide_group + year_factor,
    family = stats::binomial(),
    data = period_records
  )
  period_vcov <- sandwich::vcovCL(
    period_model,
    cluster = period_records$source_record_id,
    type = "HC1"
  )
  coefficient_names <- paste0(
    "homicide_group",
    filicide_detailed_subtype_ids
  )
  beta <- stats::coef(period_model)[coefficient_names]
  standard_error <- sqrt(diag(period_vcov))[coefficient_names]
  p_value <- 2 * stats::pnorm(
    abs(beta / standard_error),
    lower.tail = FALSE
  )

  data.frame(
    period = period_label,
    start_year = min(period_years),
    end_year = max(period_years),
    homicide_group = filicide_detailed_subtype_ids,
    homicide_group_label = unname(
      model_group_labels[filicide_detailed_subtype_ids]
    ),
    adjusted_odds_ratio_vs_other_homicide = exp(unname(beta)),
    cluster_robust_ci_lower = exp(
      unname(beta - stats::qnorm(0.975) * standard_error)
    ),
    cluster_robust_ci_upper = exp(
      unname(beta + stats::qnorm(0.975) * standard_error)
    ),
    p_value = unname(p_value),
    holm_adjusted_p_value = stats::p.adjust(
      unname(p_value),
      method = "holm"
    ),
    stringsAsFactors = FALSE
  )
}

period_specific_results <- rbind(
  estimate_period_specific_odds_ratios("2016-2020", 2016:2020),
  estimate_period_specific_odds_ratios("2021-2025", 2021:2025)
)

# Cook's distance is strongly affected by the much larger reference cells.
# Leave-one-year-out estimates provide a direct influence sensitivity check.
grouped_group_coefficient_names <- paste0(
  "homicide_group_factor",
  filicide_detailed_subtype_ids
)
full_grouped_odds_ratios <- exp(
  stats::coef(grouped_model)[grouped_group_coefficient_names]
)
leave_one_year_out_results <- do.call(
  rbind,
  lapply(analysis_years, function(omitted_year) {
    reduced_data <- annual_model_counts[
      annual_model_counts$year != omitted_year,
      ,
      drop = FALSE
    ]
    reduced_data$year_factor <- factor(reduced_data$year)
    reduced_model <- stats::glm(
      cbind(female_offender_records, male_offender_records) ~
        homicide_group_factor + year_factor,
      family = stats::binomial(),
      data = reduced_data
    )
    reduced_odds_ratios <- exp(
      stats::coef(reduced_model)[grouped_group_coefficient_names]
    )
    data.frame(
      omitted_year = omitted_year,
      homicide_group = filicide_detailed_subtype_ids,
      homicide_group_label = unname(
        model_group_labels[filicide_detailed_subtype_ids]
      ),
      full_model_adjusted_odds_ratio = unname(full_grouped_odds_ratios),
      leave_one_year_out_adjusted_odds_ratio = unname(reduced_odds_ratios),
      percent_change_from_full_model = 100 * unname(
        reduced_odds_ratios / full_grouped_odds_ratios - 1
      ),
      stringsAsFactors = FALSE
    )
  })
)
maximum_leave_one_year_out_change <- max(
  abs(leave_one_year_out_results$percent_change_from_full_model)
)

# Missing sex is not treated as a third substantive outcome. Its pattern and
# worst-case female-proportion bounds are reported separately.
missingness_results <- pooled_model_counts[c(
  "homicide_group",
  "homicide_group_label",
  "male_offender_records",
  "female_offender_records",
  "unknown_or_not_specified_offender_records",
  "known_sex_offender_records",
  "total_offender_records"
)]
missingness_results$unknown_proportion_all_records <- with(
  missingness_results,
  unknown_or_not_specified_offender_records / total_offender_records
)
missingness_results$female_proportion_lower_bound <- with(
  missingness_results,
  female_offender_records / total_offender_records
)
missingness_results$female_proportion_upper_bound <- with(
  missingness_results,
  (female_offender_records + unknown_or_not_specified_offender_records) /
    total_offender_records
)

knownness_table <- with(
  missingness_results,
  cbind(
    known_sex = known_sex_offender_records,
    unknown_or_not_specified = unknown_or_not_specified_offender_records
  )
)
knownness_test <- suppressWarnings(stats::chisq.test(
  knownness_table,
  correct = FALSE
))

zero_annual_known_sex_cells <- sum(
  annual_model_counts$male_offender_records == 0L |
    annual_model_counts$female_offender_records == 0L
)

model_diagnostics <- data.frame(
  diagnostic = c(
    "Primary logistic model converged (1=yes)",
    "Known-sex offender records in primary model",
    "SHR source-record clusters",
    "Annual group cells with zero male or zero female records",
    "Pearson overdispersion ratio",
    "Grouped-model Pearson goodness-of-fit",
    "Cells above Cook's-distance screening threshold",
    "Maximum Cook's distance",
    "Maximum absolute OR change in leave-one-year-out analysis (%)",
    "Robust omnibus homicide-group Wald test",
    "Robust group-by-transition-period Wald test",
    "Known-versus-unknown sex Pearson test",
    "Minimum expected count in knownness test"
  ),
  estimate = c(
    as.numeric(primary_model$converged),
    nrow(known_sex_records),
    length(unique(known_sex_records$source_record_id)),
    zero_annual_known_sex_cells,
    pearson_dispersion,
    pearson_chisq,
    sum(cell_diagnostics$flag_cooks_distance),
    max(cell_diagnostics$cooks_distance),
    maximum_leave_one_year_out_change,
    omnibus_group_test[["statistic"]],
    transition_interaction_test[["statistic"]],
    unname(knownness_test$statistic),
    min(knownness_test$expected)
  ),
  degrees_of_freedom = c(
    NA,
    NA,
    NA,
    NA,
    pearson_degrees_of_freedom,
    pearson_degrees_of_freedom,
    NA,
    NA,
    NA,
    omnibus_group_test[["degrees_of_freedom"]],
    transition_interaction_test[["degrees_of_freedom"]],
    unname(knownness_test$parameter),
    NA
  ),
  p_value = c(
    NA,
    NA,
    NA,
    NA,
    NA,
    pearson_gof_p,
    NA,
    NA,
    NA,
    omnibus_group_test[["p_value"]],
    transition_interaction_test[["p_value"]],
    knownness_test$p.value,
    NA
  ),
  stringsAsFactors = FALSE
)

utils::write.csv(model_diagnostics, model_diagnostics_csv, row.names = FALSE, na = "")
utils::write.csv(cell_diagnostics, cell_diagnostics_csv, row.names = FALSE, na = "")
utils::write.csv(missingness_results, missingness_csv, row.names = FALSE, na = "")
utils::write.csv(
  period_specific_results,
  period_sensitivity_csv,
  row.names = FALSE,
  na = ""
)
utils::write.csv(
  leave_one_year_out_results,
  leave_one_year_out_csv,
  row.names = FALSE,
  na = ""
)
```

The group-by-period test is statistically significant when its $p$-value is
below the chosen threshold. In that event, the pooled model should be described
as a ten-year average and the following period-specific estimates should be
reported as a sensitivity analysis.

```{r period-sensitivity-table, echo=FALSE}
period_sensitivity_display <- data.frame(
  Period = period_specific_results$period,
  Group = period_specific_results$homicide_group_label,
  `Adjusted OR (95% cluster-robust CI)` = sprintf(
    "%.2f (%.2f, %.2f)",
    period_specific_results$adjusted_odds_ratio_vs_other_homicide,
    period_specific_results$cluster_robust_ci_lower,
    period_specific_results$cluster_robust_ci_upper
  ),
  check.names = FALSE
)

knitr::kable(
  period_sensitivity_display,
  align = c("l", "l", "l"),
  caption = "Sensitivity estimates before and after the 2021 reporting transition"
)
```

```{r diagnostics-table, echo=FALSE}
diagnostic_display <- data.frame(
  Diagnostic = c(
    "Primary model converged",
    "Annual cells with a zero male or female count",
    "Pearson overdispersion ratio",
    "Pearson goodness-of-fit p",
    "Cells above Cook's-distance screen",
    "Largest absolute leave-one-year-out OR change",
    "Age group x 2021 transition p",
    "Known vs. unknown sex p"
  ),
  Result = c(
    ifelse(primary_model$converged, "Yes", "No"),
    zero_annual_known_sex_cells,
    sprintf("%.3f", pearson_dispersion),
    format.pval(pearson_gof_p, digits = 3, eps = 0.001),
    sum(cell_diagnostics$flag_cooks_distance),
    sprintf("%.1f%%", maximum_leave_one_year_out_change),
    format.pval(
      transition_interaction_test[["p_value"]],
      digits = 3,
      eps = 0.001
    ),
    format.pval(knownness_test$p.value, digits = 3, eps = 0.001)
  ),
  check.names = FALSE
)

knitr::kable(
  diagnostic_display,
  align = c("l", "r"),
  caption = "Pre-specified model checks"
)
```

```{r diagnostic-plots, echo=FALSE, fig.width=10, fig.height=4.5}
old_graphics_parameters <- graphics::par(no.readonly = TRUE)
graphics::par(mfrow = c(1, 2), mar = c(4.2, 4.2, 2.5, 1))

graphics::plot(
  cell_diagnostics$fitted_female_proportion,
  cell_diagnostics$standardized_pearson_residual,
  pch = 19,
  col = "#2C7FB8",
  xlab = "Fitted female proportion",
  ylab = "Standardized Pearson residual",
  main = "Residuals versus fitted"
)
graphics::abline(h = c(-2, 0, 2), lty = c(2, 1, 2), col = "grey50")

graphics::plot(
  cell_diagnostics$cooks_distance,
  type = "h",
  lwd = 2,
  col = "#D95F0E",
  xlab = "Year-by-group cell",
  ylab = "Cook's distance",
  main = "Cell influence"
)
graphics::abline(h = influence_threshold, lty = 2, col = "grey40")
graphics::par(old_graphics_parameters)
```

A normal Q--Q plot and a Shapiro--Wilk normality test are intentionally not
used. A binomial logistic model does not assume normally distributed outcomes
or normally distributed residuals. The relevant diagnostics are sparse cells,
convergence or separation, extra-binomial variation, influence, temporal
heterogeneity, and dependence among offender records.

The annual disjoint input, inferential estimates, model diagnostics,
cell-level diagnostics, and missingness sensitivity table are written to the
`output/` directory during knitting.

## Figures

The primary figure separates three quantities that should not be conflated:
the female proportion, the female:male ratio within each group, and the
year-adjusted odds ratio comparing that composition with other homicide. The
two odds axes are logarithmic so equal distances represent equal
multiplicative changes. All error bars are 95% confidence intervals; the
proportion intervals are Wilson intervals and the adjusted odds-ratio
intervals use source-record-clustered standard errors.

```{r figure-setup, include=FALSE}
figure_packages <- c("ggplot2", "scales", "patchwork")
missing_figure_packages <- figure_packages[!vapply(
  figure_packages,
  requireNamespace,
  quietly = TRUE,
  FUN.VALUE = logical(1L)
)]
if (length(missing_figure_packages) > 0L) {
  stop(
    sprintf(
      "Install the required figure package(s) and knit again: %s",
      paste(missing_figure_packages, collapse = ", ")
    )
  )
}

figure_directory <- file.path(output_directory, "figures")
dir.create(figure_directory, recursive = TRUE, showWarnings = FALSE)

age_figure_labels <- c(
  filicide_birth_through_6_days = "Birth-6 days",
  filicide_days_7_through_364 = "7-364 days",
  filicide_years_1_through_4 = "1-4 years",
  filicide_years_5_through_9 = "5-9 years",
  filicide_years_10_through_14 = "10-14 years",
  filicide_years_15_through_17 = "15-17 years"
)
age_figure_levels <- rev(unname(age_figure_labels))

publication_theme <- function(base_size = 12) {
  ggplot2::theme_minimal(base_size = base_size) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      panel.grid.major.y = ggplot2::element_blank(),
      plot.title.position = "plot",
      plot.title = ggplot2::element_text(face = "bold"),
      plot.subtitle = ggplot2::element_text(color = "grey30"),
      plot.caption = ggplot2::element_text(
        color = "grey35",
        hjust = 0,
        size = ggplot2::rel(0.85)
      ),
      axis.title.y = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(face = "bold"),
      legend.position = "top"
    )
}

save_figure_pair <- function(plot, filename_stub, width, height) {
  ggplot2::ggsave(
    filename = file.path(figure_directory, paste0(filename_stub, ".png")),
    plot = plot,
    width = width,
    height = height,
    units = "in",
    dpi = 320,
    bg = "white"
  )
  ggplot2::ggsave(
    filename = file.path(figure_directory, paste0(filename_stub, ".svg")),
    plot = plot,
    device = grDevices::svg,
    width = width,
    height = height,
    units = "in",
    bg = "white"
  )
}
```

### Primary comparison

```{r primary-comparison-figure, echo=FALSE, fig.width=17, fig.height=7.3, out.width="100%", fig.alt="Three-panel chart showing female offender proportions, actual within-group female-to-male ratios, and adjusted odds ratios versus other homicide across six filicide victim-age groups."}
primary_plot_data <- inferential_results[
  inferential_results$homicide_group %in% filicide_detailed_subtype_ids,
  ,
  drop = FALSE
]
primary_plot_data$age_label <- factor(
  unname(age_figure_labels[primary_plot_data$homicide_group]),
  levels = age_figure_levels
)
primary_plot_data$raw_odds_ci_lower <- with(
  primary_plot_data,
  female_proportion_ci_lower / (1 - female_proportion_ci_lower)
)
primary_plot_data$raw_odds_ci_upper <- with(
  primary_plot_data,
  female_proportion_ci_upper / (1 - female_proportion_ci_upper)
)
other_homicide_female_proportion <- inferential_results$female_proportion_known_sex[
  inferential_results$homicide_group == "other_homicide"
]
other_homicide_female_to_male_odds <- inferential_results$female_to_male_odds[
  inferential_results$homicide_group == "other_homicide"
]

primary_proportion_plot <- ggplot2::ggplot(
  primary_plot_data,
  ggplot2::aes(y = age_label, x = female_proportion_known_sex)
) +
  ggplot2::geom_vline(
    xintercept = other_homicide_female_proportion,
    linetype = "dashed",
    color = "grey35",
    linewidth = 0.6
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      xmin = female_proportion_ci_lower,
      xmax = female_proportion_ci_upper
    ),
    orientation = "y",
    width = 0.16,
    linewidth = 0.7,
    color = "#005A8D"
  ) +
  ggplot2::geom_point(size = 3.1, color = "#0072B2") +
  ggplot2::geom_text(
    ggplot2::aes(
      x = 0.88,
      label = scales::percent(female_proportion_known_sex, accuracy = 0.1)
    ),
    hjust = 1,
    size = 3.4,
    color = "#003F5C"
  ) +
  ggplot2::scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    breaks = seq(0, 0.8, by = 0.2),
    limits = c(0, 0.9),
    expand = ggplot2::expansion(mult = c(0, 0.02))
  ) +
  ggplot2::labs(
    title = "A. Female proportion",
    subtitle = sprintf(
      "Dashed line: all other homicide (%.1f%%)",
      100 * other_homicide_female_proportion
    ),
    x = "Female share among known-sex offender records"
  ) +
  publication_theme()

primary_raw_odds_plot <- ggplot2::ggplot(
  primary_plot_data,
  ggplot2::aes(y = age_label, x = female_to_male_odds)
) +
  ggplot2::geom_vline(
    xintercept = other_homicide_female_to_male_odds,
    linetype = "dotted",
    color = "grey35",
    linewidth = 0.7
  ) +
  ggplot2::geom_vline(
    xintercept = 1,
    linetype = "solid",
    color = "grey50",
    linewidth = 0.6
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      xmin = raw_odds_ci_lower,
      xmax = raw_odds_ci_upper
    ),
    orientation = "y",
    width = 0.16,
    linewidth = 0.7,
    color = "#5B4B8A"
  ) +
  ggplot2::geom_point(size = 3.1, color = "#7B61A8") +
  ggplot2::geom_text(
    ggplot2::aes(
      x = 4.7,
      label = sprintf("%.2f:1", female_to_male_odds)
    ),
    hjust = 1,
    size = 3.3,
    color = "#47336E"
  ) +
  ggplot2::scale_x_log10(
    breaks = c(0.125, 0.25, 0.5, 1, 2),
    labels = c("0.125", "0.25", "0.5", "1", "2"),
    limits = c(0.1, 5),
    expand = ggplot2::expansion(mult = c(0, 0.02))
  ) +
  ggplot2::labs(
    title = "B. Actual female:male ratio",
    subtitle = sprintf(
      "1 = parity; dotted line: other homicide (%.3f:1)",
      other_homicide_female_to_male_odds
    ),
    x = "Female records per male record (log scale)"
  ) +
  publication_theme() +
  ggplot2::theme(
    axis.text.y = ggplot2::element_blank(),
    axis.ticks.y = ggplot2::element_blank()
  )

primary_or_plot <- ggplot2::ggplot(
  primary_plot_data,
  ggplot2::aes(
    y = age_label,
    x = adjusted_odds_ratio_vs_other_homicide
  )
) +
  ggplot2::geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "grey35",
    linewidth = 0.6
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      xmin = cluster_robust_ci_lower,
      xmax = cluster_robust_ci_upper
    ),
    orientation = "y",
    width = 0.16,
    linewidth = 0.7,
    color = "#A13D00"
  ) +
  ggplot2::geom_point(size = 3.1, color = "#D55E00") +
  ggplot2::geom_text(
    ggplot2::aes(
      x = 42,
      label = sprintf("%.2f\u00d7", adjusted_odds_ratio_vs_other_homicide)
    ),
    hjust = 1,
    size = 3.4,
    color = "#713000"
  ) +
  ggplot2::scale_x_log10(
    breaks = c(1, 2, 5, 10, 20),
    labels = c("1", "2", "5", "10", "20"),
    limits = c(0.9, 45),
    expand = ggplot2::expansion(mult = c(0, 0.02))
  ) +
  ggplot2::labs(
    title = "C. Adjusted comparative odds ratio",
    subtitle = "Relative to other homicide; not a within-group ratio",
    x = "Odds ratio (log scale)"
  ) +
  publication_theme() +
  ggplot2::theme(
    axis.text.y = ggplot2::element_blank(),
    axis.ticks.y = ggplot2::element_blank()
  )

primary_comparison_figure <- patchwork::wrap_plots(
  primary_proportion_plot,
  primary_raw_odds_plot,
  primary_or_plot,
  ncol = 3,
  widths = c(1.12, 0.98, 0.9)
) +
  patchwork::plot_annotation(
    title = sprintf(
      "Female representation in filicide relative to other homicide, %d-%d",
      analysis_start_year,
      analysis_end_year
    ),
    caption = paste0(
      "Birth-6 days illustrates the distinction: 61 female / 31 male = ",
      "1.97:1 within the group, whereas the adjusted comparative OR is ",
      "15.80 because other homicide is strongly male-dominated ",
      "(16,606 female / 130,486 male = 0.127:1).\n",
      "Points show estimates; bars show 95% confidence intervals."
    ),
    theme = ggplot2::theme(
      plot.title = ggplot2::element_text(face = "bold", size = 16),
      plot.caption = ggplot2::element_text(
        color = "grey35",
        hjust = 0,
        size = 10
      )
    )
  )

save_figure_pair(
  primary_comparison_figure,
  "figure_1_primary_comparison_2016_2025",
  width = 17,
  height = 7.3
)
primary_comparison_figure
```

### Reporting-period sensitivity

```{r period-sensitivity-figure, echo=FALSE, fig.width=10, fig.height=6.5, out.width="90%", fig.alt="Forest plot comparing adjusted odds ratios for six filicide victim-age groups in 2016 through 2020 and 2021 through 2025."}
period_plot_data <- period_specific_results
period_plot_data$age_label <- factor(
  unname(age_figure_labels[period_plot_data$homicide_group]),
  levels = age_figure_levels
)
period_position <- ggplot2::position_dodge(width = 0.55)

period_sensitivity_figure <- ggplot2::ggplot(
  period_plot_data,
  ggplot2::aes(
    y = age_label,
    x = adjusted_odds_ratio_vs_other_homicide,
    color = period,
    shape = period
  )
) +
  ggplot2::geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "grey35",
    linewidth = 0.6
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      xmin = cluster_robust_ci_lower,
      xmax = cluster_robust_ci_upper
    ),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7,
    position = period_position
  ) +
  ggplot2::geom_point(size = 3, position = period_position) +
  ggplot2::scale_x_log10(
    breaks = c(1, 2, 5, 10, 20),
    labels = c("1", "2", "5", "10", "20"),
    limits = c(0.9, 30)
  ) +
  ggplot2::scale_color_manual(
    values = c("2016-2020" = "#0072B2", "2021-2025" = "#D55E00")
  ) +
  ggplot2::scale_shape_manual(
    values = c("2016-2020" = 16, "2021-2025" = 17)
  ) +
  ggplot2::labs(
    title = "Adjusted estimates differ across the 2021 reporting transition",
    subtitle = sprintf(
      "Joint group-by-period test: p = %.3f",
      transition_interaction_test[["p_value"]]
    ),
    x = "Female:male odds ratio versus all other homicide (log scale)",
    color = NULL,
    shape = NULL,
    caption = paste(
      "Models adjust for year within period.",
      "Bars are 95% source-record-clustered confidence intervals."
    )
  ) +
  publication_theme()

save_figure_pair(
  period_sensitivity_figure,
  "figure_2_period_sensitivity_2016_2025",
  width = 10,
  height = 6.5
)
period_sensitivity_figure
```

### Annual female proportions

```{r annual-proportion-figure, echo=FALSE, fig.width=13, fig.height=7, out.width="100%", fig.alt="Six faceted annual point and confidence-interval plots of female offender proportions by filicide victim-age group, with the annual all-other-homicide proportion overlaid as a reference."}
annual_proportion_intervals <- t(vapply(
  seq_len(nrow(annual_model_counts)),
  function(row_number) {
    wilson_interval(
      annual_model_counts$female_offender_records[[row_number]],
      annual_model_counts$known_sex_offender_records[[row_number]]
    )
  },
  numeric(2L)
))
annual_plot_data <- annual_model_counts
annual_plot_data$female_proportion_ci_lower <-
  annual_proportion_intervals[, "lower"]
annual_plot_data$female_proportion_ci_upper <-
  annual_proportion_intervals[, "upper"]
annual_reference_data <- annual_plot_data[
  annual_plot_data$homicide_group == "other_homicide",
  ,
  drop = FALSE
]
annual_plot_data <- annual_plot_data[
  annual_plot_data$homicide_group %in% filicide_detailed_subtype_ids,
  ,
  drop = FALSE
]
annual_plot_data$facet_label <- factor(
  age_figure_labels[annual_plot_data$homicide_group],
  levels = unname(age_figure_labels)
)

annual_proportion_figure <- ggplot2::ggplot(
  annual_plot_data,
  ggplot2::aes(x = year, y = female_proportion_known_sex)
) +
  ggplot2::geom_vline(
    xintercept = 2020.5,
    linetype = "dashed",
    color = "grey50",
    linewidth = 0.55
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      ymin = female_proportion_ci_lower,
      ymax = female_proportion_ci_upper
    ),
    width = 0.12,
    linewidth = 0.55,
    color = "#4C78A8"
  ) +
  ggplot2::geom_line(
    data = annual_reference_data,
    ggplot2::aes(x = year, y = female_proportion_known_sex),
    inherit.aes = FALSE,
    linewidth = 0.7,
    linetype = "dotted",
    color = "grey30"
  ) +
  ggplot2::geom_line(linewidth = 0.55, color = "#2C5F8A") +
  ggplot2::geom_point(size = 2, color = "#0072B2") +
  ggplot2::facet_wrap(~ facet_label, ncol = 3) +
  ggplot2::scale_x_continuous(
    breaks = c(2016, 2018, 2020, 2022, 2024),
    minor_breaks = NULL
  ) +
  ggplot2::scale_y_continuous(
    labels = scales::label_percent(accuracy = 1),
    breaks = seq(0, 1, by = 0.25),
    limits = c(0, 1)
  ) +
  ggplot2::labs(
    title = "Annual female proportions reveal sampling variability and period change",
    subtitle = paste(
      "Vertical dashed line separates reporting periods;",
      "grey dotted series is all other homicide"
    ),
    x = "Data year",
    y = "Female share among known-sex offender records",
    caption = "Points are annual observed proportions; bars are 95% Wilson intervals."
  ) +
  publication_theme() +
  ggplot2::theme(
    axis.title.y = ggplot2::element_text(angle = 90, vjust = 1),
    axis.text.x = ggplot2::element_text(angle = 45, hjust = 1)
  )

save_figure_pair(
  annual_proportion_figure,
  "figure_3_annual_female_proportion_2016_2025",
  width = 13,
  height = 7
)
annual_proportion_figure
```

### Unknown-sex records

```{r unknown-sex-figure, echo=FALSE, fig.width=10, fig.height=6.5, out.width="90%", fig.alt="Horizontal bar chart showing that unknown or unspecified offender sex is much more prevalent in all other homicide records than in the six classified filicide groups."}
unknown_plot_data <- missingness_results
unknown_plot_data$plot_label <- factor(
  c(
    "other_homicide" = "All other homicide",
    age_figure_labels
  )[unknown_plot_data$homicide_group],
  levels = rev(c("All other homicide", unname(age_figure_labels)))
)
unknown_plot_data$group_type <- ifelse(
  unknown_plot_data$homicide_group == "other_homicide",
  "All other homicide",
  "Filicide age group"
)

unknown_sex_figure <- ggplot2::ggplot(
  unknown_plot_data,
  ggplot2::aes(
    y = plot_label,
    x = unknown_proportion_all_records,
    fill = group_type
  )
) +
  ggplot2::geom_col(width = 0.62) +
  ggplot2::geom_text(
    ggplot2::aes(
      label = scales::percent(unknown_proportion_all_records, accuracy = 0.1)
    ),
    hjust = -0.12,
    size = 3.6
  ) +
  ggplot2::scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(0, 0.29),
    breaks = seq(0, 0.25, by = 0.05),
    expand = ggplot2::expansion(mult = c(0, 0))
  ) +
  ggplot2::scale_fill_manual(
    values = c(
      "All other homicide" = "#D55E00",
      "Filicide age group" = "#0072B2"
    )
  ) +
  ggplot2::labs(
    title = "Missing offender sex differs sharply across analytic groups",
    subtitle = sprintf(
      "Known-versus-unknown sex test: p %s",
      format.pval(knownness_test$p.value, digits = 3, eps = 0.001)
    ),
    x = "Unknown or not-specified share of offender records",
    fill = NULL,
    caption = paste(
      "Unknown sex is analyzed as missing outcome information, not as a",
      "substantive third category."
    )
  ) +
  publication_theme() +
  ggplot2::theme(legend.position = "none")

save_figure_pair(
  unknown_sex_figure,
  "figure_4_unknown_sex_2016_2025",
  width = 10,
  height = 6.5
)
unknown_sex_figure
```
